---
title: "Hand on ex 3 - Chap 5 Geographical Segmentation with Spatially Constrained Clustering Techniques"
editor: visual
---

# Overview

# Step 1: Loading the required packages

We load the required packages as follow:

```{r}
pacman::p_load(sf, rgdal, spdep,
               tidyverse,
               tmap,
               corrplot, ggpubr, heatmaply,
               cluster, 
               factoextra, NbClust, psych,
               here)
```

Next we generate the link to the file:

```{r}
shapefile_path <- here("data", "dataMyanmar", "geospatial")
shapefile_path
```

Now we load the shapefile: \
(question: why never set the CRS here ?)

```{r}
shan_sf <- st_read(dsn = shapefile_path, 
                   layer = "myanmar_township_boundaries") %>% 
           filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)"))
```

We check the CRS as follow:

```{r}
st_crs(shan_sf)
```

We also get a view of the data:

```{r}
glimpse(shan_sf)
```

# Step 2: Import Aspatial Data

Create the link:

```{r}
aspatial_path <- here("data", "dataMyanmar", "aspatial", "Shan-ICT.csv")
aspatial_path
```

Load the CSV file using read_csv() from readr package:

```{r}
ict <- read_csv (aspatial_path)
```

Note: the file is saved in the tibble file format. Which is different from sf format.

------------------------------------------------------------------------

Next we get the summary statistics:

```{r}
summary(ict)
```
