---
title: "Hand on ex 1 - Geodata wrangling"
editor: visual
---

## Overview Ex 1

### Step 1: Load packages

```{r}
pacman::p_load(sf,tidyverse)
```

### Step 2: Import Geo spatial Data

Import shapefile into R as a polygon feature dataframe :

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

Import shapefile into R as a line feature dataframe :

```{r}
cyclingpath <- st_read(dsn = "data/geospatial", layer = "CyclingPath")
```

Import KML into R as a Point format:

```{r}
preschookl <- st_read("data/geospatial/pre-schools-location-kml.kml")
```

### Step 3: Check content of the data frame 

This function provides a simple overview of the dataframe:

```{r}
st_geometry(mpsz)
```

This function gives a detailed breakdown of the columns and the data within the column:

```{r}
glimpse(mpsz)
```

Here we will take a detailed look at the first 5 rows of the dataset:

```{r}
head(mpsz, n=5)
```

### Step 4: Plotting the data

Below command will plot a map for each of the columns:

```{r}
plot(mpsz)
```

In order to plot the whole map only, we use the following:

```{r}
plot(st_geometry(mpsz))
```

Also we can choose to plot based on a specific column or "feature":

```{r}
plot(mpsz["PLN_AREA_N"])
```

Note\*\*\*\
plot() is only meant for a quick look, for cartographic quality, we should use other packages such as tamp.

### Step 5: Working with projections 
