---
title: "hands_on_ex_1_geo_data_wrangling"
editor: visual
---

## Overview S

In this hands-on exercise, I learned how to import cool data and wrangle them using the proper R packages.

## Getting started

Load packages:

```{r}
pacman::p_load(sf,tidyverse)
```

Import the geospatial data:

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

Import the polyline feature data in shape form:

```{r}
cyclingpath <- st_read(dsn = "data/geospatial", 
                         layer = "CyclingPath")
```

Import GIS data in kml format:

```{r}
preschool <- st_read("data/geospatial/pre-schools-location-kml.kml")
```

Checking the Content of A Simple Feature Data Frame

```{r}
st_geometry(mpsz)
```

Working with glimpse

```{r}
glimpse(mpsz)
```

Working with head:

```{r}
head(mpsz, n=5)  
```

Plotting the geospatial data:

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

### Assigning EPSG code to a simple feature data frame

```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{r}
st_crs(mpsz3414)
```

### Transforming the projection of preschool from wgs84 to svy21

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

# Import Aspatial data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

```{r}
list(listings) 
```

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
```

# Geoprocessing with sf package

Buffering

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

```{r}
sum(buffer_cycling$AREA)
```

Point-in-polygon count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

```{r}
summary(mpsz3414$`PreSch Count`)
```

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

# EDA

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

### END OF PART 1

# Choropleth Mapping with R

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

Importing data into R

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

Examine data:

```{r}
mpsz
```

Import the attribute data into R

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

Wrangle the data:

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

Joining attribute data and geospatial data

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## Plots:

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, fill = "DEPENDENCY")
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```

Data classification methods of Tmap

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

map with custom break

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

color scheme:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

map legend

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

map style:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

Cartographic Furniture

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

Multiple maps

```{r}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

facets

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)

```

arrange

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

selection criterion

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```
